Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the `eechidna` `R` package.

### 2019-03-25 04-modelling.Rmd
latex table with models

Estimation is done using generalized least squares (using the `spdep` `R` package [@spdep]).

@@

*insert table here*

>>>>

Estimation is done using feasible generalized least squares.

@@

```{r printmodels, eval = F}
# Add rho and make adjustments to this
temp_fgls <- function(my_formula, my_data, sp_weights) {
  # Spatial weights matrix
  w_mat <- listw2mat(sp_weights)

  # Get OLS residuals
  ols_model <- lm(my_formula, my_data)
  my_res <- ols_model$residuals

  # Solve for rho
  res_model <- lm(my_res ~ w_mat%*%my_res)
  rho <- res_model$coefficients[2]
  rho_df <- data.frame(estimate = rho, se = summary(res_model)$coefficients[2,2], p = summary(res_model)$coefficients[2,4])

  # Transform data for GLS
  trans_mat <- diag(nrow(my_data)) - rho*w_mat
  gls_data <- data.frame(LNP_Percent = trans_mat %*% my_data$LNP_Percent,
    Intercept = trans_mat %*% rep(1,nrow(my_data))) %>% 
    bind_cols(as.data.frame(trans_mat %*% as.matrix(my_data %>% select(-LNP_Percent))))

  # GLS model
  my_formula <- formula(paste0(my_formula,  " - 1"))
  gls_model <- gls(my_formula, gls_data)

  return(gls_model)
}

# Models
full_formula = "LNP_Percent ~ ."

# 2016
glsmod16 <- temp_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2016") %>% select(-c(DivisionNm, year)),
  sp_weights = sp_weights_16)

# 2013
glsmod13 <- temp_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2013") %>% select(-c(DivisionNm, year)),
  sp_weights = sp_weights_13)

# 2010
glsmod10 <- temp_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2010") %>% select(-c(DivisionNm, year)),
  sp_weights = sp_weights_10)

# 2007
glsmod07 <- temp_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2007") %>% select(-c(DivisionNm, year)),
  sp_weights = sp_weights_07)

# 2004
glsmod04 <- temp_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2004") %>% select(-c(DivisionNm, year)),
  sp_weights = sp_weights_04)

# 2001
glsmod01 <- temp_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2001") %>% select(-c(DivisionNm, year)),
  sp_weights = sp_weights_01)

# Stargazer
stargazer::stargazer(glsmod16, glsmod13, glsmod10, glsmod07, glsmod04, glsmod01, header = FALSE, 
  column.sep.width = "1pt", 
  title = "Estimated model using the final set of predictor variables for each of the six elections.",
  dep.var.labels = "Two-party preferred vote in favor of the Liberal party", 
  font.size = "scriptsize", df = FALSE, digits = 2, 
  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
  notes = c("Estimated coefficients for variable named in column one shown for", "election year indicated by column heading, with estimated standard", "deviation for each coefficient shown below in parenthesis. Overall", "summary measures for each regression equation are provided in the", "bottom panel."),
  notes.align = "l")
```

\begin{table}[!htbp] \centering 
  \caption{Estimated model using the final set of predictor variables for each of the six elections.} 
  \label{} 
\scriptsize 
\begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{6}{c}{Two-party preferred vote in favor of the Liberal party} \\ 
 & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
 $\rho$ & 0.48$^{***}$ & 0.29$^{*}$ & 0.15 & $-$0.04 & 0.15 & 0.44$^{***}$ \\ 
  & (0.16) & (0.16) & (0.17) & (0.19) & (0.17) & (0.15) \\ 
  & & & & & & \\ 
 AusCitizen & $-$1.53 & $-$0.87 & 1.00 & 0.67 & $-$2.83 & $-$2.77 \\ 
  & (2.95) & (2.90) & (2.93) & (2.57) & (2.63) & (2.51) \\ 
  & & & & & & \\ 
 Born\_Asia & $-$2.13 & $-$2.56 & $-$7.58$^{***}$ & $-$1.37 & 0.74 & 2.63 \\ 
  & (2.20) & (2.58) & (2.68) & (1.96) & (2.55) & (3.03) \\ 
  & & & & & & \\ 
 Born\_MidEast & $-$0.82 & 1.70 & 1.79 & 2.92$^{*}$ & 3.22$^{*}$ & 1.92 \\ 
  & (1.56) & (1.54) & (1.54) & (1.51) & (1.78) & (1.53) \\ 
  & & & & & & \\ 
 Born\_SE\_Europe & $-$1.63 & $-$1.25 & $-$1.72 & $-$0.20 & 0.04 & 1.02 \\ 
  & (0.99) & (1.15) & (1.34) & (1.17) & (2.05) & (2.34) \\ 
  & & & & & & \\ 
 Born\_UK & $-$1.37 & $-$0.35 & $-$0.29 & 0.29 & $-$0.87 & $-$0.49 \\ 
  & (0.98) & (1.00) & (1.05) & (0.80) & (0.90) & (1.04) \\ 
  & & & & & & \\ 
 BornElsewhere & 1.74 & 4.04 & 9.78$^{**}$ & 5.01 & $-$0.51 & $-$1.08 \\ 
  & (4.52) & (4.46) & (4.47) & (3.59) & (3.67) & (3.40) \\ 
  & & & & & & \\ 
 Buddhism & 1.51 & 1.09 & 2.41$^{*}$ & 0.90 & 2.84$^{*}$ & 1.50 \\ 
  & (1.24) & (1.25) & (1.40) & (1.25) & (1.61) & (1.94) \\ 
  & & & & & & \\ 
 Christianity & 5.17$^{**}$ & 3.08 & $-$0.60 & $-$1.93 & 3.44 & 0.53 \\ 
  & (2.52) & (2.79) & (3.44) & (3.45) & (4.19) & (4.58) \\ 
  & & & & & & \\ 
 CurrentlyStudying & $-$0.09 & 1.73 & 3.28$^{***}$ & 4.03$^{***}$ & 1.69 & $-$1.14 \\ 
  & (1.13) & (1.17) & (1.21) & (1.12) & (1.09) & (1.01) \\ 
  & & & & & & \\ 
 DeFacto & $-$11.02$^{***}$ & $-$9.73$^{***}$ & $-$8.60$^{***}$ & $-$6.55$^{***}$ & $-$5.04$^{**}$ & $-$6.42$^{***}$ \\ 
  & (2.55) & (2.62) & (2.57) & (1.90) & (2.00) & (1.69) \\ 
  & & & & & & \\ 
 DiffAddress & 6.93$^{***}$ & 5.02$^{***}$ & 4.72$^{***}$ & 3.40$^{***}$ & 4.63$^{***}$ & 3.86$^{***}$ \\ 
  & (1.50) & (1.71) & (1.60) & (0.88) & (1.03) & (0.95) \\ 
  & & & & & & \\ 
 Distributive & 1.65 & 2.76$^{**}$ & 2.88$^{**}$ & 2.25$^{**}$ & 2.75$^{**}$ & 1.67 \\ 
  & (1.17) & (1.25) & (1.30) & (1.02) & (1.13) & (1.08) \\ 
  & & & & & & \\ 
 Extractive & 6.81$^{***}$ & 6.09$^{***}$ & 6.97$^{***}$ & 5.81$^{***}$ & 5.31$^{***}$ & 4.69$^{***}$ \\ 
  & (1.31) & (1.42) & (1.48) & (1.16) & (1.23) & (1.36) \\ 
  & & & & & & \\ 
 Indigenous & 4.47$^{***}$ & 6.12$^{***}$ & 5.90$^{***}$ & 6.81$^{***}$ & 5.62$^{***}$ & 3.97$^{***}$ \\ 
  & (1.38) & (1.49) & (1.49) & (1.17) & (1.35) & (1.26) \\ 
  & & & & & & \\ 
 Islam & 2.64 & 1.20 & $-$1.04 & $-$1.21 & 0.61 & 0.15 \\ 
  & (1.72) & (1.71) & (1.94) & (1.80) & (2.00) & (2.03) \\ 
  & & & & & & \\ 
 Judaism & 1.96$^{***}$ & 1.60$^{*}$ & 0.46 & 0.01 & 1.74 & 1.02 \\ 
  & (0.71) & (0.85) & (1.05) & (1.00) & (1.26) & (1.38) \\ 
  & & & & & & \\ 
 ManagerAdminClericalSales & 4.62$^{***}$ & 3.99$^{***}$ & 4.32$^{***}$ & 5.21$^{***}$ & 3.28$^{***}$ & 2.10$^{***}$ \\ 
  & (1.03) & (0.99) & (1.00) & (0.78) & (0.85) & (0.70) \\ 
  & & & & & & \\ 
 Married & $-$1.39 & 1.46 & 5.07$^{*}$ & 5.06$^{*}$ & 5.50$^{**}$ & 2.27 \\ 
  & (2.64) & (2.71) & (2.91) & (2.59) & (2.79) & (2.26) \\ 
  & & & & & & \\ 
 MedianAge & 2.94$^{*}$ & 2.28 & 2.08 & 1.86 & 3.73$^{**}$ & 1.54 \\ 
  & (1.63) & (1.88) & (2.08) & (1.65) & (1.55) & (1.30) \\ 
  & & & & & & \\ 
 NoReligion & 5.76$^{**}$ & 3.53 & 0.25 & $-$0.31 & 3.43 & 1.06 \\ 
  & (2.59) & (2.58) & (2.98) & (2.90) & (3.45) & (3.85) \\ 
  & & & & & & \\ 
 OneParent\_House & $-$0.85 & 0.11 & 0.22 & 1.66 & 1.37 & $-$1.00 \\ 
  & (1.39) & (1.46) & (1.61) & (1.38) & (1.49) & (1.34) \\ 
  & & & & & & \\ 
 OtherLanguageHome & 1.27 & $-$4.69 & $-$4.37 & $-$6.02 & $-$6.76 & $-$10.40$^{**}$ \\ 
  & (4.95) & (5.50) & (5.70) & (4.44) & (5.61) & (4.97) \\ 
  & & & & & & \\ 
 SocialServ & 2.90$^{***}$ & 0.94 & 1.91 & 1.77 & 1.89 & 2.42$^{*}$ \\ 
  & (1.10) & (1.22) & (1.35) & (1.14) & (1.27) & (1.27) \\ 
  & & & & & & \\ 
 Transformative & 4.75$^{***}$ & 3.74$^{**}$ & 3.28$^{*}$ & 5.20$^{***}$ & 5.74$^{***}$ & 2.99$^{**}$ \\ 
  & (1.41) & (1.57) & (1.74) & (1.47) & (1.59) & (1.41) \\ 
  & & & & & & \\ 
 Education & $-$3.04 & $-$0.90 & $-$1.82 & $-$0.77 & 1.94 & 0.62 \\ 
  & (2.99) & (3.04) & (3.35) & (2.89) & (2.86) & (2.23) \\ 
  & & & & & & \\ 
 FamHouseSize & $-$4.35 & $-$6.16$^{*}$ & $-$8.22$^{**}$ & $-$9.67$^{***}$ & $-$5.11$^{*}$ & $-$2.94 \\ 
  & (2.86) & (3.15) & (3.28) & (2.66) & (2.78) & (2.27) \\ 
  & & & & & & \\ 
 PropertyOwned & 1.76 & 0.68 & $-$1.17 & 0.78 & $-$0.92 & $-$0.76 \\ 
  & (1.38) & (1.43) & (1.49) & (1.09) & (1.21) & (1.19) \\ 
  & & & & & & \\ 
 RentLoanPrice & $-$0.70 & 1.01 & 2.12 & $-$0.54 & $-$1.54 & $-$0.31 \\ 
  & (2.02) & (1.96) & (1.96) & (1.48) & (1.68) & (1.50) \\ 
  & & & & & & \\ 
 Incomes & 8.46$^{***}$ & 4.19 & 2.73 & 4.36$^{*}$ & 2.95 & 2.99$^{*}$ \\ 
  & (2.57) & (2.68) & (3.06) & (2.56) & (2.60) & (1.76) \\ 
  & & & & & & \\ 
 Unemployment & 0.36 & $-$1.22 & $-$2.42 & $-$2.88$^{**}$ & $-$5.09$^{***}$ & $-$3.00$^{**}$ \\ 
  & (1.42) & (1.54) & (1.73) & (1.41) & (1.51) & (1.28) \\ 
  & & & & & & \\ 
 Constant & 50.45$^{***}$ & 53.48$^{***}$ & 49.92$^{***}$ & 47.32$^{***}$ & 52.63$^{***}$ & 50.78$^{***}$ \\ 
  & (0.74) & (0.56) & (0.51) & (0.35) & (0.45) & (0.67) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
Log Likelihood & $-$399.52 & $-$405.41 & $-$418.08 & $-$398.73 & $-$403.34 & $-$398.33 \\ 
Akaike Inf. Crit. & 863.03 & 874.83 & 900.16 & 861.46 & 870.68 & 860.67 \\ 
Bayesian Inf. Crit. & 951.96 & 963.76 & 989.10 & 950.40 & 959.61 & 949.60 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{6}{l}{Estimated coefficients for variable named in column one shown for} \\ 
 & \multicolumn{6}{l}{election year indicated by column heading, with estimated standard} \\ 
 & \multicolumn{6}{l}{deviation for each coefficient shown below in parenthesis. Overall} \\ 
 & \multicolumn{6}{l}{summary measures for each regression equation are provided in the} \\ 
 & \multicolumn{6}{l}{bottom panel.} \\
\end{tabular} 
\end{table}
